---
title: 'Assignmment 1'
author: "Daniel Alonso"
date: "November 20th, 2020"
output: 'pdf_document'
---

Importing libraries

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
library(Rcpp)
library(microbenchmark)
```

## Teacher example (this code is NOT mine)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
my_knn_R = function(X, X0, y){
  # X data matrix with input attributes
  # y response variable values of instances in X  
  # X0 vector of input attributes for prediction
  
  nrows = nrow(X)
  ncols = ncol(X)
  
  # One of the instances is going to be the closest one:
  # closest_distance: it is the distance , min_output
  closest_distance = 99999999
  closest_output = -1
  closest_neighbor = -1
  
  for(i in 1:nrows){
    
    distance = 0
    for(j in 1:ncols){
      difference = X[i,j]-X0[j]
      distance = distance + difference * difference
    }
    
    distance = sqrt(distance)
    
    if(distance < closest_distance){
      closest_distance = distance
      closest_output = y[i]
      closest_neighbor = i
    }
  }
  closest_output
}  
```

## Testing teacher example (This code is NOT mine)

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# X contains the inputs as a matrix of real numbers
data("iris")
# X contains the input attributes (excluding the class)
X <- iris[,-5]
# y contains the response variable (named medv, a numeric value)
y <- iris[,5]
# From dataframe to matrix
X <- as.matrix(X)
# From factor to integer
y <- as.integer(y)
# This is the point we want to predict
X0 <- c(5.80, 3.00, 4.35, 1.30)
# Using my_knn and FNN:knn to predict point X0
# Using the same number of neighbors, it should be similar (k=1)
my_knn_R(X, X0, y)
library(FNN)
FNN::knn(X, matrix(X0, nrow = 1), y, k=1)
```


\newpage

## Translating the teacher code into C++ into an Rccp function

```{r, echo=TRUE, warning=FALSE, message=FALSE}
cppFunction('
int knn_1(NumericMatrix X, NumericVector X0, NumericVector y) {
    int nrows = X.nrow();
    int ncols = X.ncol();

    double closest_distance = 99999999;
    double closest_output = -1;
    double closest_neighbor = -1;
    double difference = 0;

    int i;
    int j;

    for (i = 0; i < nrows; i++) {
        
        double distance = 0;
        for (j = 0; j < ncols; j++) {
            difference = X(i,j) - X0(j);
            distance = distance + difference * difference;
        }

        distance = sqrt(distance);

        if (distance < closest_distance) {
            closest_distance = distance;
            closest_output = y(i);
            closest_neighbor = i;
        }
    }
    return closest_output;
}')
```

## Testing Rcpp translation

```{r, echo=TRUE, warning=FALSE, message=FALSE}
knn_1(X, X0, y)
library(FNN)
FNN::knn(X, matrix(X0, nrow = 1), y, k=1)
```

## Benchmarking differences in runtime between R version and Rcpp version

### R version
```{r, echo=TRUE, warning=FALSE, message=FALSE}
microbenchmark(my_knn_R(X, X0, y))
```

We can see that our mean runtime for the R version is more than 800 microseconds


### FNN version

```{r, echo=TRUE, warning=FALSE, message=FALSE}
microbenchmark(FNN::knn(X, matrix(X0, nrow = 1), y, k=1))
```

We can see that our mean runtime for the Rcpp version is of under 250 microseconds

### Rcpp version

```{r, echo=TRUE, warning=FALSE, message=FALSE}
microbenchmark(knn_1(X, X0, y))
```

We can see that our mean runtime for the Rcpp version is of under 14 microseconds

