---
title: 'Assignmment 1'
author: "Daniel Alonso"
date: "November 20th, 2020"
output: 'pdf_document'
---

Importing libraries

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
library(Rcpp)
library(microbenchmark)
library(foreach)
```

# Exercise 1

## Class example (this code is NOT mine)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
my_knn_R = function(X, X0, y){
  # X data matrix with input attributes
  # y response variable values of instances in X  
  # X0 vector of input attributes for prediction
  
  nrows = nrow(X)
  ncols = ncol(X)
  
  # One of the instances is going to be the closest one:
  # closest_distance: it is the distance , min_output
  closest_distance = 99999999
  closest_output = -1
  closest_neighbor = -1
  
  for (i in 1:nrows) {

    distance = 0
    for (j in 1:ncols) {
      difference = X[i,j]-X0[j]
      distance = distance + difference * difference
    }
    
    distance = sqrt(distance)
    
    if (distance < closest_distance) {
      closest_distance = distance
      closest_output = y[i]
      closest_neighbor = i
    }
  }
  closest_output
}  
```

## Testing class example (This code is NOT mine)

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# X contains the inputs as a matrix of real numbers
data("iris")
# X contains the input attributes (excluding the class)
X <- iris[,-5]
# y contains the response variable (named medv, a numeric value)
y <- iris[,5]
# From dataframe to matrix
X <- as.matrix(X)
# From factor to integer
y <- as.integer(y)
# This is the point we want to predict
X0 <- c(5.80, 3.00, 4.35, 1.30)
# Using my_knn and FNN:knn to predict point X0
# Using the same number of neighbors, it should be similar (k=1)
my_knn_R(X, X0, y)
library(FNN)
FNN::knn(X, matrix(X0, nrow = 1), y, k=1)
```


\newpage

## Translating the teacher's code into C++ into an Rccp function

```{r, echo=TRUE, warning=FALSE, message=FALSE}
cppFunction('
int knn_1(NumericMatrix X, NumericVector X0, NumericVector y) {
    int nrows = X.nrow();
    int ncols = X.ncol();

    double closest_distance = 99999999;
    double closest_output = -1;
    double closest_neighbor = -1;
    double difference = 0;

    int i;
    int j;

    for (i = 0; i < nrows; i++) {
        
        double distance = 0;
        for (j = 0; j < ncols; j++) {
            difference = X(i,j) - X0(j);
            distance = distance + difference * difference;
        }

        distance = sqrt(distance);

        if (distance < closest_distance) {
            closest_distance = distance;
            closest_output = y(i);
            closest_neighbor = i;
        }
    }
    return closest_output;
}')
```

## Testing Rcpp translation

```{r, echo=TRUE, warning=FALSE, message=FALSE}
knn_1(X, X0, y)
library(FNN)
FNN::knn(X, matrix(X0, nrow = 1), y, k=1)
```

## Benchmarking differences in runtime between R version and Rcpp version

### R version
```{r, echo=TRUE, warning=FALSE, message=FALSE}
microbenchmark(my_knn_R(X, X0, y))
```

We can see that our mean runtime for the R version is more than 800 microseconds


### FNN version

```{r, echo=TRUE, warning=FALSE, message=FALSE}
microbenchmark(FNN::knn(X, matrix(X0, nrow = 1), y, k=1))
```

We can see that our mean runtime for the Rcpp version is of under 250 microseconds

### Rcpp version

```{r, echo=TRUE, warning=FALSE, message=FALSE}
microbenchmark(knn_1(X, X0, y))
```

We can see that our mean runtime for the Rcpp version is of under 14 microseconds

# Exercise 2

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
knn_more = function(X, X0, y, K){
  # X data matrix with input attributes
  # y response variable values of instances in X  
  # X0 vector of input attributes for prediction
  
  nrows = nrow(X)
  ncols = ncol(X)
  
  # One of the instances is going to be the closest one:
  # closest_distance: it is the distance , min_output
  distances = c()
  closest_distance = 1e99
  closest_neighbor = -1
  closest_classif = -1
  
  # get distances
  for (i in 1:nrows) {

    distance = 0
    for (j in 1:ncols) {
      difference = X[i,j]-X0[j]
      distance = distance + difference * difference
    }
    
    distance = sqrt(distance)

    # add distance to vector
    distances = c(distances, distance)

    if (distance < closest_distance) {
      closest_distance = distance
      closest_classif = y[i]
      closest_neighbor = i
    }
  }

  # eliminating closest distance
  NN_distances = c(closest_distance)
  NN_classif = c(closest_classif)
  distances[closest_neighbor] = 1e99
  distances = unname(distances)

  # We already got the closest so remove one from K
  K = K - 1
  
  # because we can't sort, we just manually pull out the minimum value K times
  # by subtracting each distance to the previous closest distance
  for (i in 1:K) {

    # placeholder variables for loop
    diff = 0
    min_diff = 1e99
    index = 0

    # calculate diffs between distances and closest distance
    # the lowest is saved in placeholder variable min_diff
    # then the index is saved in the index variable
    for (idx in 1:nrows) {
      diff = distances[idx] - NN_distances[i]
      if (diff < min_diff) {
        min_diff = diff
        index = idx
      }
    }

    # add the corresponding distance to NN distances
    # add the corresponding classif to NN classif
    NN_distances = c(NN_distances, distances[index])
    NN_classif = c(NN_classif, y[index])
    distances[index] = 1e99
  }

  # different classifications
  classifs = unique(y)

  # loop through classifications to count
  cnts = matrix(rep(0,6), nrow=length(classifs), byrow=TRUE)
  cnts[,1] = classifs;
  for (g in NN_classif) {
    cnts[g,2] = cnts[g,2] + 1
  }

  # check if there's identical counts
  count_vector = cnts[,2]
  group_normally = 0
  if (K <= length(classifs)) {
    if (length(count_vector[count_vector == max(count_vector)]) > 1) {
      for (i in 1:K) {
        if (NN_distances[i] == min(NN_distances)) {
          group = NN_classif[i]
        }
      }
    } else {
      group_normally = 1
    } 
  } 
  else {
    group_normally = 1
  }
  
  # select maximum value
  if (group_normally == 1) {
    group = 0
    for (i in cnts[,1]) {
      if (count_vector[i] == max(count_vector)) {
        group = i
      }
    }
  }

  group
}  
test <- knn_more(X, X0, y, 3)
test
```

## Translating the teacher code into C++ into an Rccp function

```{r, echo=TRUE, warning=FALSE, message=FALSE}
cppFunction('
int knn_more(NumericMatrix X, NumericVector X0, NumericVector y, int K) {
    int nrows = X.nrow();
    int ncols = X.ncol();

    NumericVector distances(nrows);
    NumericVector NN_distances(K);
    NumericVector NN_classif(K);
    double closest_distance = 99999999999999999;
    double closest_output = -1;
    double closest_neighbor = -1;
    double difference;

    int i;
    int j;
    

    for (i = 0; i < nrows; i++) {
        
        double distance = 0;
        for (j = 0; j < ncols; j++) {
            difference = X(i,j) - X0(j);
            distance = distance + difference * difference;
        }

        distance = sqrt(distance);
        distances[i] = distance;


        if (distance < closest_distance) {
            closest_distance = distance;
            closest_output = y(i);
            closest_neighbor = i;
        }
    }

    K = K - 1;
    NN_distances(0) = closest_distance;
    NN_classif(0) = closest_output;
    distances(closest_neighbor) = 99999999999999999;
    
    int idx;
    for (i = 0; i < K; i++) {
      double diff = 0;
      double min_diff = 99999999999999999;
      int index = 0;
      for (idx = 0; idx < nrows; idx++) {
        diff = distances(idx) - NN_distances(i);
        if (diff < min_diff) {
          min_diff = diff;
          index = idx;
        }
      }
      NN_distances(i+1) = distances(index);
      NN_classif(i+1) = y(index);
      distances(index) = 99999999999999999;
    }
    
    NumericVector classifs(unique(y).size());
    for (i = 0; i < unique(y).size(); i++) {
      classifs[i] = i+1;
    }

    NumericMatrix cnt(classifs.size(), 2);
    for (i = 0; i < classifs.size(); i++) {
      cnt(i,0) = classifs(i);
      cnt(i,1) = 0;
    }
    for (i = 0; i < NN_classif.size(); i++) {
      cnt(NN_classif(i)-1,1) = cnt(NN_classif(i)-1,1) + 1;
    }

    NumericVector count_vector = cnt(_,1);
    NumericVector maxes = count_vector[count_vector == max(count_vector)];
    int group = 0;
    int group_normally = 0;
    int maxes_size = maxes.size();
    if (K <= classifs.size()) {
      if (maxes_size > 1) {
        for (i = 0; i < K; i++) {
          if (NN_distances(i) == min(NN_distances)) {
            group = NN_classif(i);
          }
        }
      } else {
        group_normally = 1;
      }
    } else {
      group_normally = 1;
    }
    
    if (group_normally == 1) {
      for (i = 0; i < classifs.size(); i++) {
        if (count_vector(i) == max(count_vector)) {
          group = i+1;
        }
      }
    }

  return group;
}')
test <- knn_more(X, X0, y, 3)
test
```